\section{Related Work}
\label{sec:related}

\noindent\textbf{Pose estimation and body reconstruction.}
Recovering the configuration of a human body from a single image is a mature problem.
In two dimensions, pose estimators already localize keypoints with high accuracy.
Early architectures fuse multi-scale features~\cite{newell2016stacked,xiao2018simple} and associate keypoints across people~\cite{cao2019openpose}.
Later designs preserve high-resolution features~\cite{sun2019deep} or adopt transformer backbones~\cite{xu2022vitpose}.
Beyond 2D joints, parametric models recover 3D pose and shape.
SMPL defines the standard representation~\cite{loper2015smpl}.
Around this representation, regressors estimate pose and shape from one image~\cite{kanazawa2018end,kolotouros2019spin,li2022cliff,goel2023humans}.
SAM 3D Body extends this line to full-body recovery under the MHR representation~\cite{yang2026sam3dbody}.
In all of these methods, geometry is the final product.
We instead treat reconstruction as annotation.
Dense outputs become interpretable measurements that ground language supervision.

\noindent\textbf{Action quality and pose correction.}
Action quality assessment scores how well an action is performed~\cite{lei2019survey}.
Methods exist for cross-action scoring~\cite{parmar2019action} and rehabilitation exercises~\cite{liao2020deep}.
For yoga and similar exercises, Pose Tutor also identifies the joint angles behind a pose prediction~\cite{dittakavi2022posetutor}.
Such outputs say whether a pose is wrong without saying how the body is configured.
Our focus is complementary to this line of work.
The supervision we construct keeps every numeric statement traceable to an explicit, image-specific measurement.

\noindent\textbf{Vision--language models and VQA.}
Vision--language models learn statistical links between appearance and text at scale.
CLIP establishes the recipe with contrastive pretraining over image--text pairs~\cite{radford2021clip}.
Building on such representations, Flamingo and BLIP-2 couple frozen image encoders with language models~\cite{alayrac2022flamingo,li2023blip2}.
Visual instruction tuning then teaches the coupled models to follow instructions~\cite{liu2023visual,dai2023instructblip}.
Qwen2-VL adds dynamic-resolution processing~\cite{wang2024qwen2vl}.
These links, however, connect appearance only to the words that describe it.
The benchmarks behind this progress remain object-centric.
VQAv2 and GQA ask about objects, attributes, and relations whose answers are visible in the image~\cite{goyal2017vqav2,hudson2019gqa}.
A body posture, by contrast, is arbitrarily linked to its cultural or social meaning.
This link cannot be recovered from appearance statistics alone.
Nearly identical body configurations can carry substantially different meanings.
No existing benchmark grounds language in the measured configuration of the body.
YogaAlign supplies this missing supervision by pairing images with explicit, measured body geometry.

\noindent\textbf{Human pose datasets.}
Existing pose datasets do not close this gap.
COCO and MPII annotate 2D keypoints at scale~\cite{lin2014coco,andriluka2014mpii}.
For 3D, Human3.6M provides motion capture under controlled conditions~\cite{ionescu2014human36m}.
3DPW extends 3D annotation to in-the-wild video~\cite{vonMarcard2018}.
Closest to our setting, Yoga-82 provides a visually diverse benchmark of 82 asanas~\cite{verma2020yoga82}.
Its supervision, however, stops at hierarchical category labels.
No existing dataset pairs an image with measured body geometry and language grounded in that geometry.
YogaAlign builds on this diversity.
It adds reconstruction-derived measurements and language-grounded supervision.
